MSF_Groupwork-NBA_shotlog

### 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
hot hand theory - continue

#Set the value to character
df[, "SHOT_RESULT"] <- as.character(df[, "SHOT_RESULT"])

#Define 1 for baskets made and 0 for baskets missed. We need a vector to use it to our function below.
shotresult <- c()
for(i in 1:nrow(df)) {
  if(shots[i] == 'missed') shotresult <- c(shotresult, 0)
  else if(shots[i] == 'made') shotresult <- c(shotresult, 1)
}

shotresult

#Set the value to numeric/integer
df[, "SHOT_RESULT"] <- as.numeric(df[, "SHOT_RESULT"])
df[ ,'SHOT_NUMBER'] <- as.integer(df[ , 'SHOT_NUMBER'])

#Define the length of a shooting streak to be the number of consecutive baskets made until a miss occurs
#Create a function that gives us a vector with the number of streaks
number_of_streaks <- function(x) {
  x <- c(0, x, 0)
  which_zero <- which(x == 0)
  streak <- diff(which_zero) - 1
  streak
}

number_of_streaks(shotresult)

player_and_game <- function(x, y) {
   player <- df[df[ ,'player_id'] == x & df[ ,'GAME_ID1'] == y, 'FGM']
   number_of_streaks(as.vector(player))
}
player_and_game(202681, 21400681)

ourtable <- head(pttsum2,10)

list_per_player <- list()
j <- 1
for(i in ourtable$player_id) {
   streak_per_player[[j]] <- player_and_game(i, ourtable[ourtable[ ,'player_id'] == i, 'GAME_ID1'])
   j <- j + 1
}
streak_per_player

>>>>

#From 'FGM' column we have that 1 represents a basket made and 0 represents a basket missed
#Define the length of a shooting streak to be the number of consecutive baskets made until a miss occurs
#Create a function that gives us a vector with the number of total streaks(from all players and all games)
number_of_streaks <- function(x) {
  x <- c(0, x, 0)
  which_zero <- which(x == 0)
  streak <- diff(which_zero) - 1
  streak
}

number_of_streaks(df$FGM)

#Set a new data frame for the 5 players with best performances(higher points)
best_players_df <- head(pttsum2,5)

#Create a function that gives us a vector of streaks for a specific player and game
player_and_game <- function(x, y) {
   player <- df[df[ ,'player_id'] == x & df[ ,'GAME_ID1'] == y, 'FGM']
   number_of_streaks(as.vector(player))
}
#For the first player and game try:
#player_and_game('202681', '21400681')

#Create a list that gives us the streaks of each player and game (5 best performances)
streak_per_player <- list()
j <- 1
for(i in best_players_df$player_id) {
   streak_per_player[[j]] <- player_and_game(i, best_players_df[ourtable[ ,'player_id'] == i, 'GAME_ID1'])
   j <- j + 1
}
streak_per_player

#Create a data frame for the strikes of each one of the 10 players (cannot create a data frame including all of them, because each list has different length)
player1_df <- data.frame('streaks' = streak_per_player[[1]])
player2_df <- data.frame('streaks' = streak_per_player[[2]])
player3_df <- data.frame('streaks' = streak_per_player[[3]])
player4_df <- data.frame('streaks' = streak_per_player[[4]])
player5_df <- data.frame('streaks' = streak_per_player[[5]])

#Show distribution for each player's streak lengths
ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[1,'player_name'])
ggplot(data = player2_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[2,'player_name'])
ggplot(data = player3_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[3,'player_name'])
ggplot(data = player4_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[4,'player_name'])
ggplot(data = player5_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[5,'player_name'])

#Create a boxplot for each player's streak lengths
boxplot(player1_df$streaks) 
summary(player1_df$streaks) #Check that boxplot is valid
IQR(player1_df$streaks) #Check that boxplot is valid
#The typical length of a streak is 0 (median = 0) - The Interquartile Range is 1 - Streak lengths of 3 and 6 are unusually high compared to the rest of the distribution
boxplot(player2_df$streaks)
summary(player2_df$streaks)
IQR(player2_df$streaks)
#The typical length of a streak is 1 (median = 1) - The Interquartile Range is 2 - There are some high streak lengths (max = 4)
boxplot(player3_df$streaks)
summary(player3_df$streaks)
IQR(player3_df$streaks)
#The typical length of a streak is 0 (median = 0) - The Interquartile Range is 0 - Streak lengths of 3 and 13 (!!!) are unusually high compared to the rest of the distribution
boxplot(player4_df$streaks)
summary(player4_df$streaks)
IQR(player4_df$streaks)
#The typical length of a streak is 1 (median = 1) - The Interquartile Range is 2 - Streak length of 6 is unusually high compared to the rest of the distribution
boxplot(player5_df$streaks)
summary(player5_df$streaks)
IQR(player5_df$streaks)
#The typical length of a streak is 1 (median = 1) - The Interquartile Range is 2 - There are some high streak lengths (max = 5)

#We can see that the distribution of all players is right skewed and that all of them (especially player3) have some long shooting streaks. To prove/disprove the hot hand theory we need to examine the indipendence of the shots. If each shot that a player takes is independent of the next shot, then a missed/made shot will not affect the propability that the next shot will be missed/made. So, if consecutive shots are independent, we disprove hot hand theory, while if consecutive shots are dependent, we prove the hot hand theory.
#Simulation - create a vector of shots('made', 'missed') for a sample shooter
set.seed(23)
shot_results <- c('made', 'missed')
simul_shooter <- sample(shot_results, size = nrow(player1_df), replace = TRUE, prob = c(0.45, 0.55))
simul_shooter

#Set 1 for baskets made and 0 for baskets missed
simul_shooter <- as.list(simul_shooter)
for(i in 1:length(simul_shooter)) {
   if(simul_shooter[i] == 'made') simul_shooter[i] <- 1
   else if(simul_shooter[i] == 'missed') simul_shooter[i] <- 0
}
simul_shooter

#Use previous function (number_of_streaks) to calculate the streaks made by the sample shooter 
streaks_of_simulshooter <- number_of_streaks(simul_shooter)
simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)

#Comparisons of data between sample shooter and real shooters
summary(streaks_of_simulshooter)
summary(player1_df$streaks)

#Compare barplots of sample shooter and real shooter1
#If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
ggplot(data = simulshooter_df, aes(x = streaks)) + geom_bar() 
ggplot(data = player1_df, aes(x = streaks)) + geom_bar()
#
```
